---
title: Exploring the U.S. National Oceanic and Atmospheric Administration's (NOAA)
  storm database
output:
  html_document:
    keep_md: yes
  pdf_document: default
header-includes: \usepackage{color}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=TRUE, warning=FALSE}
## LIBRARIES
library(stringr)
library(dplyr, warn.conflicts = FALSE)
library(ggplot2)

```

Generated `r str_to_title(format(Sys.Date(), "%B  %e, %Y"))`


# Synopsis

This project involves exploring the U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database. This database tracks characteristics of major storms and weather events in the United States, including when and where they occur, as well as estimates of any fatalities, injuries, and property damage.

We want to answers two questions:  
1. Across the United States, which types of events (as indicated in the EVTYPE variable) are most
harmful with respect to population health?  
2. Across the United States, which types of events have the greatest economic consequences?  


# Data Processing

We are interested to study population health and economic consequences across the United States for all types of events:

**States**:  
- COUNTY: county numeric code  
- COUNTYNAME: county name   
- STATE: state code  

**Population health**:  
- FATALITIES: approx. number of deaths  
- INJURIES: approx. number of injuries

**Economic consequences**:  
- PROPDMG: approx. property damages  
- PROPDMGEXP: the units for property damage value  
- CROPDMG: approx. crop damages   
- CROPDMGEXP: the units for crop damage value  

**Events**:  
- EVTYPE: weather event (Tornados, Wind, Snow, etc..)  


## Pre-analysis

The fist step is load the data into R and do a preanalysis: 

```{r, echo=FALSE, warning=FALSE}
## CLEAN OBJECTS FROM WORKSPACE
rm(list=ls())

## CHANGE WORKING DIRECTORY
setwd("C:/Users/ABC/Cursos/Data Scientists - Johns Hopkins University/WorkingDirectory_Curso/datasciencecoursera_repository/Course 5/Project_2/")

# TO PRINT BIG NUMBERS
options(scipen = 20)

```

```{r, echo=TRUE, cache = TRUE, warning=FALSE}
## UNZIP AND LOAD DATA
if(!file.exists("storm_data.csv"))
    unzip("storm_data.zip")
storm_data <- read.csv("storm_data.csv", sep=",", header=TRUE, stringsAsFactors = FALSE)


## NEW NUMBER OF TYPE OF EVENTS
length(sort(unique(storm_data$EVTYPE)))
head(sort(table(storm_data$EVTYPE), decreasing = TRUE))

## CLEAN ALL TYPE OF EVENTS' NAMES
# - All letters to uppercase
storm_data$EVTYPE <- toupper(storm_data$EVTYPE)
# - Refactory the name:
storm_data$EVTYPE <-  gsub("EROSIN", "EROSION",
                           gsub("TORNDAO", "TORNADO", 
                                gsub("AVALANCE", "AVALANCHE", 
                                     gsub("SUMMARY:", "SUMMARY", 
                                          gsub("SUMMARY OF", "SUMMARY", 
                                               gsub("LIGHTING", "LIGHTNING",
                                                    storm_data$EVTYPE)))))) 

storm_data$EVENT_ALIAS <- "OTHER"
storm_data$EVENT_ALIAS[grep("HAIL", storm_data$EVTYPE, ignore.case = TRUE)] <- "HAIL" 
storm_data$EVENT_ALIAS[grep("WIND", storm_data$EVTYPE, ignore.case = TRUE)] <- "WIND" 
storm_data$EVENT_ALIAS[grep("FLOOD", storm_data$EVTYPE, ignore.case = TRUE)] <- "FLOOD" 
storm_data$EVENT_ALIAS[grep("TORNADO", storm_data$EVTYPE, ignore.case = TRUE)] <- "TORNADO" 
storm_data$EVENT_ALIAS[grep("LIGHTNING", storm_data$EVTYPE, ignore.case = TRUE)] <- "LIGHTNING" 
storm_data$EVENT_ALIAS[grep("SNOW", storm_data$EVTYPE, ignore.case = TRUE)] <- "SNOW"  
storm_data$EVENT_ALIAS[grep("STORM", storm_data$EVTYPE, ignore.case = TRUE)] <- "STORM"  
storm_data$EVENT_ALIAS[grep("RAIN", storm_data$EVTYPE, ignore.case = TRUE)] <- "RAIN" 
storm_data$EVENT_ALIAS[grep("WINTER", storm_data$EVTYPE, ignore.case = TRUE)] <- "WINTER" 
storm_data$EVENT_ALIAS[grep("FIRE", storm_data$EVTYPE, ignore.case = TRUE)] <- "FIRE" 


## NEW NUMBER OF TYPE OF EVENTS
length(sort(unique(storm_data$EVENT_ALIAS)))
sort(table(storm_data$EVENT_ALIAS), decreasing = TRUE)

```


## Types of events most harmful with respect to population health
We need to aggreate the number of casualties by the event type.
```{r, echo=TRUE, warning=FALSE}
## FATALITIES/INJURIES BY EVENT TYPE 
storm_data_by_group <- group_by(storm_data, EVENT_ALIAS)
casualties <- summarize(storm_data_by_group,  
                        fatalities=sum(FATALITIES), 
                        injuries =sum(INJURIES),
                        total=sum(FATALITIES)+sum(INJURIES) 
) %>% arrange(desc(total))


## MOST HARMFULL
casualties[1,]

```


## Types of events have the greatest economic consequences
We need to look at the property damage and crop damage data reportings.   
- The property damage is represented with two fields: a number **PROPDMG** in dollars and the exponent **PROPDMGEXP**.  
- The crop damage is represented using two fields: **CROPDMG** in dollars and the exponent **CROPDMGEXP**.   

The first step in the analysis is to calculate the property and crop damage for each event.

```{r, echo=TRUE, warning=FALSE}

## EXPONENT EQUIVALENCE
sort(unique(storm_data$PROPDMGEXP))
sort(unique(storm_data$CROPDMGEXP))
exp_equivalence <- function(e){
    # K or k: thousand dollars (1000) (k)
    # M or m: million dollars (1000000) (m)
    # B or b: billion dollars (1000000000) (b)
    # the rest would be consider as dollars (d)
    if (e %in% c("h", "H"))
        return(2)
    else if (e %in% c("k", "K"))
        return(3)
    else if (e %in% c("m", "M"))
        return(6)
    else if (e %in% c("b", "B"))
        return(9)
    else if (!is.na(as.numeric(e))) 
        return(as.numeric(e))
    else if (e %in% c("", "-", "?", "+"))
        return(0)
    else {
        stop("Invalid exponent value.")
    }
}

## CALCULATE THE PROPERTY AND CROP DAMAGE
storm_data$PROPERTY <- storm_data$PROPDMG * (10 ^ exp_equivalence(storm_data$PROPDMGEXP))
storm_data$CROP <- storm_data$CROPDMG * (10 ^ exp_equivalence(storm_data$CROPDMGEXP))


## PROPERTY/CROP BY EVENT TYPE 
storm_data_by_group <- group_by(storm_data, EVENT_ALIAS)
damages <- summarize(storm_data_by_group,  
                     property=sum(PROPERTY), 
                     crop =sum(CROP),
                     total=sum(PROPERTY)+sum(CROP) 
) %>% arrange(desc(total))


## MOST HARMFULL
damages[1,]

```

# Results

After this analysis, we concluded that:

1- The **\textcolor{blue}{`r casualties[1, ]$EVENT_ALIAS`}** is the event type that cause most number of deaths and injuries among all event types, with a total of `r casualties[1, ]$total` (fatalities `r casualties[1, ]$fatalities` and injuries `r casualties[1, ]$injuries`) 

```{r, echo=TRUE, warning=FALSE}
## PLOT TOP 5 

ggplot(casualties[1:5,], aes(x=EVENT_ALIAS, y=total)) + geom_bar(stat="identity") +
    xlab("") + ylab("") +
    ggtitle("TOP 5 most harmful with respect to population health") +
    theme_bw() +
    theme(
        plot.title = element_text(size=12,lineheight=.7, face="bold", hjust = 0.5, vjust = -1.0, 
                                  color= "black"),
        axis.text.x = element_text(size=7, color= "black"),
        legend.position="none"
    )  


```

2- The **\textcolor{blue}{`r damages[1, ]$EVENT_ALIAS`}** is the event type that have the greatest economic consequences among all event types, with a total of `r damages[1, ]$total` (property damages `r damages[1, ]$property` and crop damages `r damages[1, ]$crop`) 

```{r, echo=TRUE, warning=FALSE}
## PLOT TOP 5 

ggplot(damages[1:5,], aes(x=EVENT_ALIAS, y=total)) + geom_bar(stat="identity") +
    xlab("") + ylab("") +
    ggtitle("TOP 5 most have the greatest economic consequences") +
    theme_bw() +
    theme(
        plot.title = element_text(size=12,lineheight=.7, face="bold", hjust = 0.5, vjust = -1.0, 
                                  color= "black"),
        axis.text.x = element_text(size=7, color= "black"),
        legend.position="none"
    )  

```
