---
title: "Reproducible Research: Peer Assessment 1"
output: 
    html_document:
        keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
```{r, echo=TRUE}
library(ggplot2)
library(lubridate,warn.conflicts = FALSE)
```


## Loading and preprocessing the data

```{r, echo=TRUE}

# CLEAN OBJECTS FROM WORKSPACE
rm(list=ls())

# WORKING DIRECTORY
setwd("C:/Users/ABC/Cursos/Data Scientists - Johns Hopkins University/WorkingDirectory_Curso/datasciencecoursera_repository/Course 5/Project_1/")

# LOAD DATA
unzip("activity.zip")
df_activity <- read.csv("activity.csv", sep=",", stringsAsFactors = FALSE, header = TRUE)

# TRANSFORM TO DATE
df_activity$date <- as.factor(as.Date(df_activity$date, format = "%Y-%m-%d"))

```



## What is mean total number of steps taken per day?

```{r, echo=TRUE}
## HISTOGRAM OF TOTAL NUMBER OF STEPS TAKEN EACH DAY 
total_steps <- with(df_activity, tapply(steps, date, FUN = sum, na.rm = TRUE))
qplot(total_steps, binwidth = 1000, ylab = "Frecuency", xlab="Number of steps per day")


## MEAN AND MEDIAM OF THE TOTAL NUMBER OF STEPS TAKEN PER DAY
mean(total_steps)
median(total_steps) 

```  


```{r, echo=TRUE}
## BARPLOT OF TOTAL OF STEPS PER DAY
total_steps <- aggregate(steps ~ date, df_activity[!is.na(df_activity$steps),], sum)
ggplot(data=total_steps, aes(x=date,y=steps)) + 
    geom_bar(stat="identity") + 
    ggtitle("Total steps per day") + xlab(" ") + ylab("Steps") +
    theme_bw() +
    theme(
        plot.title = element_text(size=12,lineheight=.7, face="bold", hjust = 0.5, vjust = -1.0, 
                                  color= "black"),
        axis.text.x = element_text(size=7, color= "black", angle=90, hjust = 1),
        legend.position="none"
    )  

```


## What is the average daily activity pattern?

```{r, echo=TRUE}
## AVERAGE DAILY OF STEPS TAKEN EACH INTERVAL
daily_average <- aggregate(steps ~ interval, df_activity[!is.na(df_activity$steps),], mean)

ggplot(daily_average, aes(x=interval, y=steps)) + geom_line() +
    xlab("5-minute interval") + ylab("Average number of steps") + 
    ggtitle("Time Series Plot")

```

```{r, echo=TRUE}
max_interval <- daily_average[which.max(daily_average$step),]
```

The **5-minute interval that contains the maximum number of steps** is: `r max_interval$interval` with `r max_interval$steps` mean of steps



## Imputing missing values

```{r, echo=TRUE}
## TOTAL OF MISSING VALUES
summary(df_activity)
table(is.na(df_activity$steps))
total_na <- sum(is.na(df_activity$steps))

## FILLING IN ALL OF THE MISSING VALUES
# - Strategy: reemplace NA by the mean for that 5 minute interval
df_activity_withoutNA <- df_activity
intervals_na <- unique(df_activity[is.na(df_activity$steps),"interval"])
for(i in intervals_na){
    df_activity_withoutNA[is.na(df_activity_withoutNA$steps) & df_activity_withoutNA$interval==i, ]$steps <- daily_average[daily_average$interval==i,]$steps
}

## HISTOGRAM OF TOTAL NUMBER OF STEPS TAKEN EACH DAY 
total_steps <- with(df_activity_withoutNA, tapply(steps, date, FUN = sum, na.rm = TRUE))
qplot(total_steps, binwidth = 1000, ylab = "Frecuency", xlab="Number of steps per day")

## MEAN AND MEDIAM OF THE TOTAL NUMBER OF STEPS TAKEN PER DAY
mean(total_steps)
median(total_steps) 

```  



## Are there differences in activity patterns between weekdays and weekends?

```{r, echo=TRUE}
## DETERMINE IS WEEKDAY OR WEEKEND DAY
df_activity_withoutNA$day <- "weekday"
df_activity_withoutNA[wday(df_activity_withoutNA$date, label=TRUE) %in% c("Sun","Sat"), ]$day <- "weekend"
table(df_activity$day)

## AVERAGE DAILY OF STEPS TAKEN EACH INTERVAL
daily_sum <- aggregate(steps ~ interval+day, df_activity_withoutNA[!is.na(df_activity_withoutNA$steps),], mean)

ggplot(daily_sum, aes(x=interval, y=steps)) + 
    geom_line() + facet_grid(day ~ .) + 
    xlab("5-minute interval") + ylab("Average number of steps") + 
    ggtitle("Time Series Plot")

```
