---
title: "Machine Learning - Project: Prediction of classe"
output:
html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).


### Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.


### Goal

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.



# Preliminary Work

An overall pseudo-random number generator seed was set at 1234 for all code. In order to reproduce the results below, the same seed should be used. 

```{r }
# - Reproduceability
set.seed(1234)

# - Load libraries
library(ggplot2)
library(caret)
library(randomForest)

# - Load Data
input_training <- read.csv(url("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"),
                           na.strings=c("NA","#DIV/0!",""))
input_testing <- read.csv(url("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"),
                          na.strings=c("NA","#DIV/0!",""))
table(input_training$classe)
```
The variable "classe" is a factor variable with 5 levels.


### Cleaning the input data

We need return a clean input data for used as training data set
```{r }
#summary(input_training)

# - Exclude variables that cannot be used for prediction: identifier, timestamp, and window data
input_training <- input_training[, - grep("name|timestamp|window|X", colnames(input_training), value=F) ]

# - Remove variables with at least 60% of NA
cs <- colSums(sapply(input_training, is.na))
input_training <- input_training[, (cs / nrow(input_training) < 0.6)]

# - Removing zero covariates
nzv <- nearZeroVar(input_training, saveMetrics = TRUE)
input_training <- input_training[, !nzv$nzv ]
```

### Working with training and testing data
```{r }
# - Create a series of testing/training partitions
inTrain <- createDataPartition(input_training$classe, p=0.7, list=FALSE)
training <- input_training[inTrain,]
testing <- input_training[-inTrain,]

# - Built a Model and Predict on the testing set
modFit <- randomForest(classe ~. , data=training)
prediction <- predict(modFit, testing)
confusionMatrix(prediction, testing$classe)

```
The result is a model with good prediction values.

### Test Cases
Now is moment to use "test cases" and predict the classe
```{r }
prediction <- predict(modFit, input_testing)
prediction

```



